\textbf{SPECFEM3D}\\ \textbf{Cartesian}\\ \textbf{User Manual}

### 2022-11-12 doc/USER_MANUAL/03_mesh_generation.tex
updates manual (minimum period estimate)

\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in geographical coordinates, and
needs to be \texttt{.true.} when your model is specified in Cartesian
coordinates. \noun{UTM projection zone in which your simulation region
resides.}
%
\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
of the topography and of the interfaces between the different layers
of the model, if any. The number of spectral elements in the vertical
direction within each layer is also defined in this file.
%
\item [{\texttt{NEX\_XI}}] The number of spectral elements along one side of the
block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
defined below. Based upon benchmarks against semi-analytical discrete
wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
that a $\nexxi=288$ run is accurate to a shortest period of roughly
2~s. Therefore, since accuracy is determined by the number of grid
points per shortest wavelength, for any particular value of $\nexxi$
the simulation will be accurate to a shortest period determined by
\begin{equation}
\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
\end{equation}
 The number of grid points in each orthogonal direction of the reference
element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
by \texttt{NGLLX} in the \texttt{setup/constants.h} file. We generally use
$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
elements. We suggest not to change this value.
%
\item [{\texttt{NEX\_ETA}}] The number of spectral elements along the other side
of the block. This number \textit{must} be 8~$\times$~a multiple
of $\nproceta$ defined below.
%
\item [{\texttt{NPROC\_XI}}] The number of processors or slices along one side
of the block (see Figure~\ref{fig:For-parallel-computing}); we must
have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
integer.
%
\item [{\texttt{NPROC\_ETA}}] The number of processors or slices along the other
side of the block; we must have $\nexeta=8\times c\times\nproceta$,
where $c\ge1$ is a positive integer.
%
\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
want a perfectly regular mesh or \texttt{.false.} if you want to add
doubling horizontal layers to coarsen the mesh. In this case, you
also need to provide additional information by setting up the next
three parameters.
%
\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
Must be set at least to \texttt{1} if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}.
Multiple mesh doublings can be chosen, for which each an \texttt{NZ\_DOUBLING\_**} entry must be given.
By default, we only provide two possible entries in the \texttt{Mesh\_Par\_file}.
For higher numbers of doubling layers, additional entries must be added.
\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}).
\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}
and if \texttt{NDOUBLINGS} is set to \texttt{2}). Doubling layers must be at least \texttt{2} layers apart.
The layer count starts from the bottom layer.
More entries must be listed by the user if \texttt{NDOUBLINGS} is larger than \texttt{2}.
\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
for subsequent viewing. Turning the flag on generates files in the
\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
for a discussion of mesh viewing features.
\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
viewing.
\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
by the mesher will be written. Generally one uses a directory on the
local disk of the compute nodes, although on some machines these partitions
are written on a parallel (global) file system (see also the earlier
discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).

>>>>

\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in geographical coordinates, and
needs to be \texttt{.true.} when your model is specified in Cartesian
coordinates. \noun{UTM projection zone in which your simulation region
resides.}
%
\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
of the topography and of the interfaces between the different layers
of the model, if any. The number of spectral elements in the vertical
direction within each layer is also defined in this file.
%
\item [{\texttt{NEX\_XI}}] The number of spectral elements along one side of the
block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
defined below. Based upon benchmarks against semi-analytical discrete
wavenumber synthetic seismograms, \cite{KoLiTrSuStSh04} determined
that a $\nexxi=288$ run for a 516 km x 507 km mesh
is accurate to a shortest period of roughly
2~s. Therefore, since accuracy is determined by the number of grid
points per shortest wavelength, for any particular value of $\nexxi$
the simulation with a mesh extent $\Delta$ (in km)
will be accurate approximately to a shortest period determined by
\begin{equation}
\mbox{shortest period (s)}=2. \times \frac{288}{\nexxi} \times \frac{\Delta}{500km}\label{eq:shortest_period}
\end{equation}
This estimate also depends on the velocity model. Therefore, the binary \texttt{xgenerate\_databases}
calculates a minimum period resolved by the chosen mesh discretization and velocity model,
output into the file \texttt{OUTPUT\_FILES/output\_generate\_databases.txt}.

 The number of grid points in each orthogonal direction of the reference
element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
by \texttt{NGLLX} in the \texttt{setup/constants.h} file. We generally use
$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
elements. We suggest this value as it represents a fair trade-off between element deformations to adapt to
topographic variations and numerical accuracy.
%
\item [{\texttt{NEX\_ETA}}] The number of spectral elements along the other side
of the block. This number \textit{must} be 8~$\times$~a multiple
of $\nproceta$ defined below.
%
\item [{\texttt{NPROC\_XI}}] The number of processors or slices along one side
of the block (see Figure~\ref{fig:For-parallel-computing}); we must
have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
integer.
%
\item [{\texttt{NPROC\_ETA}}] The number of processors or slices along the other
side of the block; we must have $\nexeta=8\times c\times\nproceta$,
where $c\ge1$ is a positive integer.
%
\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
want a perfectly regular mesh or \texttt{.false.} if you want to add
doubling horizontal layers to coarsen the mesh. In this case, you
also need to provide additional information by setting up the next
three parameters.
%
\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
Must be set at least to \texttt{1} if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}.
Multiple mesh doublings can be chosen, for which each an \texttt{NZ\_DOUBLING\_**} entry must be given.
By default, we only provide two possible entries in the \texttt{Mesh\_Par\_file}.
For higher numbers of doubling layers, additional entries must be added.
\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}).
\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}
and if \texttt{NDOUBLINGS} is set to \texttt{2}). Doubling layers must be at least \texttt{2} layers apart.
The layer count starts from the bottom layer.
More entries must be listed by the user if \texttt{NDOUBLINGS} is larger than \texttt{2}.
\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
for subsequent viewing. Turning the flag on generates files in the
\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
for a discussion of mesh viewing features.
\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
viewing.
\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
by the mesher will be written. Generally one uses a directory on the
local disk of the compute nodes, although on some machines these partitions
are written on a parallel (global) file system (see also the earlier
discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
